Automated optimization of a course of actions in a stochastic environment

### 2016-05-31 paper.Rmd
fdsfdf
dfdf

RL emerges easily from other Machine Learning techniques as the training is done by acting. It uses training information that *evaluates* the actions taken rather than *instructs* by giving correct actions are faced repeatedly with a choice among n different options, or actions. Such setting has its roots in the *n-armed bandit problem* where after each choice you receive a numerical reward chosen from a stationary probability distribution that depends on the selected action. The objective is to maximize the expected total reward over all the episodes. The learner evolves in a particular state and records what actions lead to a maximal gain. In most cases the goal is to maximise a long term gain by taing a correct course of actions. These policies can be more or less explicit and the proposed algorithm would qualified as *off-policy*.

@@

##1.2 Recognising real world situations

@@

#3 Proposed principle

@@

\newcommand{\IND}[1][1]{\STATE\hspace{#1\algorithmicindent}}
\begin{algorithm*}
\caption{}
\begin{algorithmic}
\STATE$\text{Initialized:}$
\IND$X\text{: empty }m\text{-columns feature matrix}$
\IND$y\text{: empty target vector}$
\IND$X'\text{: empty }m\text{-columns simulated states matrix}$
\IND$w^*=w_0\text{: random weights for the regressor}$
\IND$\bm{\gamma} \in [0,1] \text{ discount rate}$
\IND$\Delta = 0 \text{: number of actions}$
\STATE$\text{Other variables:}$
\IND$f() \text{: regressor function}$
\IND$m \text{: number of features}$
\IND$p_t(x^*) \text{: stochastic process, function of the last existing chosen state}$
\IND$x_t(p_t) \in {\rm I\!R^m} \text{: current state vector}$
\IND$\Omega_t(x_t) \text{: set of possible actions}$
\IND$\epsilon \text{: randomness rate}$
\IND$s(x^*) \in {\rm I\!R} \text{: signal value triggered by the modified state}$
\STATE$\textbf{WHILE } \text{the learning process is on:}$
\IND$\text{record }x_t$
\IND$X' := \text{ empty matrix}$
\IND$\textbf{FOR } x' \text{ in } \Omega_t(x_t)\text{:}$
\IND[2]$\text{append } x' \text{ to } X'$
\IND$\text{With probability } 1-\epsilon$
\IND[2]$x^*:=\arg\max_{x' \in X'} f(w^*,X')$
\IND$\text{With probability } \epsilon$
\IND[2]$x^*:=\text{random sample }x' \in X'$
\IND$\text{append } x^* \text{ to } X$
\IND$\textbf{IF } s(x^*) \text{ != } 0 \text{ :}$
\IND[2]$\textbf{FOR } \delta \text{ in } 1\text{:}\Delta \text{:}$
\IND[3]$value = d(\gamma, \Delta, \delta, s(x^*))$
\IND[3]$\text{append } value \text{ to } y$
\IND[2]$w^* := \arg\min_{w}(l(f(w,X),y))$
\IND[2]$\Delta := 0$
\IND$\textbf{ELSE} \text{:}$
\IND[2]$\Delta \text{ += } 1$
\end{algorithmic}
\end{algorithm*}

@@

The use of this simple sensor is made possible by the use of discounted
values. Even if not every action is monitored during an episode, each step still takes a value at the end of the episodes. In that process, tuning correctly the $\gamma$ parameter is important. In the following figure, we represent two sets of actions that to respectively a gain, and a loss.

for gain and loss, display same gradient: from red to green
SHOW IMAGE TREE WITH DIFFERENT DISCOUNT FACTORS

We see that with a higher discount factor, a stronger emphasis is put on the final result of the set of actions. In some cases, $\gamma$ should be set considering that two intermediary actions can lead to both a gain and a loss.

@@

#4 Asymptotic behaviour of the algorithm

@@

##4.1 Convergence for equivalent underlying states

>>>>

RL emerges easily from other Machine Learning techniques as the training is done by acting. It uses training information that *evaluates* the actions taken rather than *instructs* by giving correct actions are faced repeatedly with a choice among n different options, or actions. Such setting has its roots in the *n-armed bandit problem* where after each choice you receive a numerical reward chosen from a stationary probability distribution that depends on the selected action. The objective is to maximize the expected total reward over all the episodes. The learner evolves in a particular state and records what actions lead to a maximal gain. In most cases the goal is to maximise a long term gain by taing a correct course of actions. These policies can be more or less explicit. I our case, the proposed algorithm would be qualified as *off-policy*.

@@

##1.2 Pattern recognition

@@

#3 Proposed algorithm

@@

\newcommand{\IND}[1][1]{\STATE\hspace{#1\algorithmicindent}}
\begin{algorithm*}
\caption{}
\begin{algorithmic}
\STATE$\text{Initialized variables:}$
\IND$X\text{: empty }m\text{-columns feature matrix}$
\IND$y\text{: empty target vector}$
\IND$w^*=w_0\text{: random weights for the regressor}$
\IND$\bm{\gamma} \in [0,1] \text{ discount rate}$
\IND$\Delta = 0 \text{: number of actions}$
\STATE$\text{Other variables:}$
\IND$f() \text{: regressor function}$
\IND$m \text{: number of features}$
\IND$p_t(x^*) \text{: stochastic process, function of the last existing chosen state}$
\IND$X'\text{: }m\text{-columns simulated states matrix}$
\IND$x_t(p_t) \in {\rm I\!R^m} \text{: current state vector}$
\IND$\Omega_t(x_t) \text{: set of possible actions}$
\IND$\epsilon \text{: randomness rate}$
\IND$s(x^*) \in {\rm I\!R} \text{: signal value potentially triggered by the modified state}$
\STATE$\textbf{WHILE } \text{the learning process is on:}$
\IND$\text{record }x_t(p_t)$
\IND$X' := \text{: empty }m\text{-columns matrix}$
\IND$\textbf{FOR } x' \text{ in } \Omega_t(x_t)\text{:}$
\IND[2]$\text{append } x' \text{ to } X'$
\IND$\text{With probability } 1-\epsilon$
\IND[2]$x^*:=\arg\max_{x' \in X'} f(w^*,X')$
\IND$\text{With probability } \epsilon$
\IND[2]$x^*:=\text{random sample }x' \in X'$
\IND$\text{append } x^* \text{ to } X$
\IND$\textbf{IF } s(x^*) \text{ != } 0 \text{ :}$
\IND[2]$\textbf{FOR } \delta \text{ in } 1\text{:}\Delta \text{:}$
\IND[3]$value = d(\gamma, \Delta, \delta, s(x^*))$
\IND[3]$\text{append } value \text{ to } y$
\IND[2]$w^* := \arg\min_{w}(l(f(w,X),y))$
\IND[2]$\Delta := 0$
\IND$\textbf{ELSE} \text{:}$
\IND[2]$\Delta \text{ += } 1$
\end{algorithmic}
\end{algorithm*}

@@

The use of this simple sensor is made possible by the use of discounted
values. Even if not every action is monitored during an episode, each step still takes a value at the end of the episodes. In that process, tuning correctly the $\gamma$ parameter is important. In the following figure, we represent a set of actions that to a gain (signal = 1 here). The discount function that we apply is $\gamma^{\Delta-\delta}s(x^*)$.

\begin{figure}
\begin{center}
\includegraphics[width=200pt]{/Users/Thomas/git/thesis/discounts.png}
\caption{Different discount rates affecting intermediary states}
\end{center}
\end{figure}

In the first case $\gamma=0.7$, in the second case $\gamma=0.3$. We see that with a larger discount factor, more emphasis is put on the final result of the set of actions. In some cases, $\gamma$ should be set considering that two intermediary actions can lead to both a gain and a loss.

@@

#4 Behaviour of the algorithm

@@

##4.1 Convergent estimates for equivalent underlying states
